Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays

### 2020-01-16 review_article.Rmd
Remove numbering

---
title: "Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
author: "Stephanie Kobakian^[Science and Engineering Faculty, Queensland University of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University.] and Jessie Roberts^[Queensland University of Technology.]"
output:
  word_document:
    toc: no
    toc_depth: '3'
  pdf_document:
    dev: jpeg
    fig_caption: yes
    number_sections: yes
    toc: no
    toc_depth: 3
geometry: margin = 1.2in
header-includes:
    - \usepackage{setspace}\doublespacing
link-citations: yes
csl: springer-vancouver.csl
fontsize: 11pt
bibliography: reviewbib.bib
yolo: yes
---

```{r setup, include=FALSE}
fig_path <- glue::glue("figs/")
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)
library(ggplot2)
library(cowplot)
library(png)
library(grid)
```

@@

# 2. Traditional approaches for cancer map displays {#ch:choropleth_maps}
<!-- 381 words -->
<!-- what is a choropleth map -->
A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI, @BCM]. Faithful rendering of the geography, when combined with an appropriate color scheme, can reveal spatial patterns among data values. Identifying and explaining spatial structures, patterns, and processes involve considering the individuals and organizing them into representable units of communities [@SAMGIS]. Early versions of choropleth maps used symbols or patterns instead of color. Choropleth maps can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA]. In epidemiology, choropleth maps are often used as a tool to study the spatial distribution of cancer incidence and mortality.

@@

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1: A selection of choropleth cancer maps from online atlases that are publicly available. Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends. These atllases are described in Table 1.", out.width='100%', fig.align='center', fig.width=12, fig.height=10}
knitr::include_graphics("figures/choropleth_grid.png")
```

Table: Table 1: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.

@@

Table: Table 2: <!--\label{tab:measures}--> Common measures for reporting cancer information.

@@

Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates. To address this, a cancer atlas often communicates uncertainty about the value of a statistic. There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modeling or simulation process. The most common measures used to present uncertainty are credible or confidence intervals (CIs). Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively. The map focuses on displaying the statistic and lacks additional space to represent the uncertainty. Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two common solutions.

## 2.3 Limitations of choropleth displays {#ch:chorolimit}

@@

<!--
```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged from sex and years 2012-2016, using four different map projections. The map projections alter the shapes and angles of the boundaries of each state.", message=FALSE, warning=FALSE, out.width='100%', fig.align='center'}
#knitr::include_graphics("figures/ggchoroCRS.png")
c1 <- png::readPNG("figures/gg3857.png")
c2 <- png::readPNG("figures/gg2163.png")
c3 <- png::readPNG("figures/gg4326.png")
c4 <- png::readPNG("figures/gg2955.png")
#knitr::include_graphics("figures/usa_grid.png")
load("figures/usa_legend.rda")

@@

```{r ggcartograms, fig.cap = "Figure 2: Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous). Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state. This provides a better sense of the extent of disease relative to the population in the country and can help ease losing information about physically small but population-dense states. Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE, out.width='100%'}
knitr::include_graphics("figures/usa_grid.png")
```

Table: Table 3: <!--\label{tab:usa}--> Maps used to present statistics for the United States of America. The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.

|Map display |Details| 
|:--------------------|:------------------------------------------------|  
|a. Contiguous| It has distorted each state's shape according to the population of the state in 2015. The state of California has become much larger because of the large population density. This draws attention to the densely populated North-East region and detracts from the less populated Mid West.|
|b. Non-contiguous| It maintains the geographic shape of the states, but the size has altered according to the population of the state in 2015. The state of California has remained closer to its original size than its surrounding states. The North-East states have remained closer to their geographical size, for Massachusetts and Connecticut. This draws attention to the densely populated North-East region and the sparse Mid West.|
|c. Dorling| Circles are used to represent each state, but the population of the state determines the size in 2015.  The North-East states remain closer to their neighbors and are slghtly displaced from their geographic location. It highlights the sparsity of the population in the Mid West by the distance between the circles at the geographic centroids.|
|d. Hexagon Tessellation| A hexagon of equal size represents each state. It is easy to contrast the neighboring states however the North-East regions have been displaced from their geographic location. It highlights the sparsity of the population in the Mid West by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|

@@

To be able to recognize the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC]. Tobler's Conformal mapping means to preserve angles locally so that the shapes of small areas on a traditional map and a cartogram would be similar.
@CBATCC presents this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.

### 3.1.2 Non-contiguous

@@

## 3.3 Geofacet
<!-- 166 words -->
Hafen [@IGF] introduces the term geofacet to describe a grid display of small plots. The arrangement of tiles mimics the geographic topology. Geofaceting has the functionality that a statistical plot can be constructed in each facet for each geographic area. A tile map can communicate only one value per region in a visualization, while geofaceting is a more flexible visualization for communication as it increases the amount of information displayed. Virtually any type of plot can be shown in the tile, allowing displays of multiple variables or values per geographic entity. Creating the layout of a geofacet is manual, but once created can be used for any data on that geographic base.

@@

## 3.4 Multivariate displays
<!-- 222 words -->

@@

# 4. Comparison and critique of alternative displays {#sec:compare}
<!-- 408 words -->

```{r auscartograms, fig.cap = "Figure 4: Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, out.width='100%', cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figures/auscartograms.png")
```

## 4.1 Neither choropleth maps or cartograms perform well for Australia

@@

<!-- Cartograms, sacrifice geography, focus on the outliers/ relationships in data space 
Cartographers were historically the creators of geographic displays, and epidemiologists utilized the maps that had been prepared in advance.
Cartograms with the population variable used to distortion the geographic map regions and boundaries are density equalizing maps, that reduce the visual prominence of large, but low-population areas [@VSSDCUC]. 
Cartogram displays incorporate the statistics and population into the design of the display,
shifting and sometimes sacrificing familiar boundaries to draw attention to large outliers in the data space or population density.
The unusual shapes are not longer familiar and may be difficult to compare. This can prevent effective comparison of the statistic. Cartograms also favor extreme statistics in the positive direction. Small extreme values are often lost due to the resulting small sized areas the do not draw user's attention.
The change in the map based on one variable provides the opportunity for bi-variate displays, using the change in area for one variable, and color for another.
The difference between the familiar map display helps highlight the impact of the disease on communities.
-->

<!-- Alternatives, remove geographic shapes, keep spatial relationship, focus on the outliers/ relationships in population space 
Alternative maps may replace the unexpected boundaries and shapes with familiar, simple shapes. This makes the spatial distribution the primary concept communicated by the display.
Areas that are geographically close will maintain connectedness in some way, but the population and the statistic will dominate.
tilemap displays may only allow one variable to be visualized if each region is given consistent map space, this occurs in geofaceting. 
Dorlings provide bi-variate displays as they use the size of the shape for one variable, and color for another.
These alternatives are especially helpful for data aggregations where administrative boundaries break populations into groups.
The experience of each community may be worth considering, as the experiences shared by the population within them may be similar due to the services and facilities they share.
This display allows a more equitable view of each community, and does not minimise those that operate on a smaller geographic scale.

Faceting provides an opportunity to contrast the overall distribution of variables across a geographic space or collection of geographic areas.
This display makes is difficult to compare specific regions.-->

Table: Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).

@@

# 5. User interaction {#ch:interacting}
<!-- 634 words -->

<!-- Interactivity and animation can add value to online atlases, by enabling additional information to be presented on-the-fly with one display. It also invites the viewer to engage with the display, and the information.

The word "interactive" is used, and perhaps, confused for many different graphics actions. At one extreme, it is synonymous with direct manipulation of the display, such as selecting an area, mousing over, or panning and zooming, using mouse action on the plot. At the other extreme, it refers to graphical user interfaces (GUI) such as selecting an item on a menu to change the variable being displayed, or checking a toggle box, or dragging a scrollbar. @RPIG provides a good overview of types of interactivity with plots. a key difference from an animation is that interactivity typically involves on-the-fly computations, and changes to the plot.

The primary interactive tools found in online cancer maps include

- drop-down menus to select features,
- variable selection, 
- tool tips, 
- zooming and panning.

The latter two would be considered to be direct manipulation on the map display, and the former two are GUI interfaces. The effects of these actions are rapid, incremental, and reversible [@DMIV]. 
-->

@@

<!-- 
XXX This part needs to go to interactivity!
One of the concerns of adding too much information to a map is the fear of cognitive overload [@mcgranaghan1993cartographic] in which the user reaches an information threshold, beyond which will not be able to make sense of the information. These concerns are not unfounded, complexity and density of representation methods appear to overwhelm novice decision makers, while experts are able to use the detail more readily when making-decisions [@cliburn2002design]. Interactivity is a design feature within modern mapping methods that can be used to incorporate additional information and complexity without overloading the user.

Interactivity enables supplementary information to be incorporated into online atlases without cluttering the display. Interactive design features, found in online cancer maps, include tooltips, drop-down menus, data selection, zooming and panning allow users to explore the map as they want more information and allow flexibility in the display [@HTLWM]. The use of these supports were found in various online cancer maps identified by @roberts2019communication. The controls for basic interactive features are often placed outside of the plot space [@TGA], thus the map image is updated/replaced as the user interacts with the controls. For example, changing the population age or other demographic variables. Some more advanced interactions include direct interactions with the plot via the use of overlaid tool tip features, very few cancer atlases involve these more complex selection tools.

Additionally, interactivity allows the user to toggle between different variables, map views or multiple realizations of possible future scenarios [@goodchild1994introduction]. Thus providing additional mechanisms for the users comprehension as well as the uncertainty of the available information (@maceachren1992visualizing; @van1994visualization).

These interactive features provide an opportunity for users to explore the additional information available. This helps users to understand and interpret the spatial distribution presented, as well as validate, explain or explore the presented statistics and their relationships to each other and/or their underlying spatial distribution. This allows relationships between spatial areas and diseases to be explored with sophistication in non-traditional but still 'cognitively accessible' ways [@TNTEA]. The interactive features of the publicly available maps identified by @roberts2019communication allow the exploration of geographic hierarchies, population distribution, statistical uncertainty, demographics and socio-economic indicators.
@TNTEA suggested LM plots as a solution to linking cartography and statistical graphics.
-->

<!-- ## 5.1 Under-utilized techniques-->

@@

<!-- It is uncommon to see linked displays utilized in online atlases, but there are several frameworks that support interactive displays of multivariate data. @TNTEA suggested LM plots as a solution to linking cartography and statistical graphics. 
Linked displays allow for direct manipulation, @RPIG uses Swayne and Klinke's definition of high level interaction that occurs via data selection or brushing tools that highlight subsets of graphical elements such as polygons or points. Selections can be made statistically through interactions which allow subsets to be selected from statistical displays, such as tables of data.
Monmonier implemented linking choropleth maps with statistical plots, adding a geographic component to scatter plots [@MCRG], the use of link allows spatial dimensions to be reserved for geographic locations of areas. The statistical variation of additional variables can be presented in other appropriate displays.
Low level interaction is more common, this occurs via the user interface or the map.
Interacting with the map returns values or information used in the graphical display without referring to, of changing, the original data source of the map.
-->

<!-- Animations -->
<!--Animations communicate a story, showcasing a critical region by defining and progressing through a small set of keyframes, or waypoints [@ESRI].
@TGA suggests the use of moving objects demands attention by taking advantage of the behaviour of the visual cognition system.
They can be produced to not tolerate or require user input, creating useful, directed descriptions of processes, such as highlighting a particular region or time period [@HTLWM].
Animations ensure all users follow the same interpretive path but creators cannot control the level of engagement.
-->

<!-- Animations can be implemented to allow users varying degrees of control.
Figure \ref{fig:animating} a gives an example of how time can be included through animating displays. This is a basic form of interactivity with animation allowing users to play, pause and rewind films, where the film is intended to be watched in one direction but the position in the animation can be controlled.
@CPISACA provide weather maps as a thoroughly developed example of animation of spatial displays to communicate information to the general public.
The movement of a weather system will follow a forecasted path, all map users can follow the animated path of the weather system across the geography over a specified period of time.
Where the message is most important, static or animated graphics allow control over the display and interpretation.
User interaction more allow clicks to progress between stages of more complex animations.
The Australian Cancer Atlas [@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) which change the display to draw user's attention to areas on the map that are relevant to the story.
This implementation of animation gives users tools to plan their own exploration.
-->

```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9}
knitr::include_graphics("figures/interacting.png")
```

```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a. Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b. A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 7}
knitr::include_graphics("figures/animating.png")
```

# 6. Conclusions
<!-- 307 words -->

<!-- summary of mapping -->
This paper provides an overview of mapping practices as commonly used for cancer atlases and recommends new approaches, such as cartograms and hexagon tile maps that should be adopted going forward. The conventional approach is the choropleth map, and it is widely used. The choropleth map suffers when there are small geographic units, as occurs in Australia where the population is concentrated on the coast, the information about the burden of cancer on those communities can be hidden. Making an inset can clarify congested regions but this breaks the viewers' attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed. The map alternatives implement tradeoffs between the familiar shapes, and the importance of the geographic areas in the context of the areas. Given the population or a cancer statistic for each area, the geographic size or shape will change. Alternative displays allow the spatial distribution of cancer data to be digested by map users.

>>>>

---
title: "Cancer Applications of Choropleth Maps, and the Potential of Cartograms and Alternative Map Displays"
author: "Stephanie Kobakian^[Science and Engineering Faculty, Queensland University of Technology.], Dianne Cook^[Faculty of Econometrics and Business Statistics, Monash University.] and Jessie Roberts^[Queensland University of Technology.]"
output:
  pdf_document:
    dev: jpeg
    fig_caption: yes
    number_sections: yes
    toc: no
    toc_depth: 3
geometry: margin = 1.2in
header-includes:
    - \usepackage{setspace}\doublespacing
link-citations: yes
csl: springer-vancouver.csl
fontsize: 11pt
bibliography: reviewbib.bib
yolo: yes
---

```{r setup, include=FALSE}
fig_path <- glue::glue("figs/")
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE,
  out.width = "\\textwidth")
library(ggplot2)
library(cowplot)
library(png)
library(grid)
```

@@

# Traditional approaches for cancer map displays {#ch:choropleth-maps}
<!-- 381 words -->
<!-- what is a choropleth map -->
A choropleth map displays the geographical distribution of data over a set of spatial units by shading areas of a map [@EI], [@BCM]. Faithful rendering of the geography, when combined with an appropriate color scheme, can reveal spatial patterns among data values. Identifying and explaining spatial structures, patterns, and processes involve considering the individuals and organizing them into representable units of communities [@SAMGIS]. Early versions of choropleth maps used symbols or patterns instead of color. Choropleth maps can be used for displaying disease data [@DMAHP], including cancer data [@CPISACA]. In epidemiology, choropleth maps are often used as a tool to study the spatial distribution of cancer incidence and mortality.

@@

```{r choropleth-grid-create, message=FALSE, warning = FALSE, fig.cap = "Figure 1: A selection of choropleth cancer maps from online atlases that are publicly available. Maps of various countries were chosen: United Kingdom, Australia, Spain, USA, Canada, and display several different colour palettes and legends. These atllases are described in Table 1.", fig.align='center', fig.width=12, fig.height=10}
knitr::include_graphics("figures/choropleth_grid.png")
```

Table 1: <!--\label{tab:choropleths}--> A selection of choropleth cancer maps from online atlases.

@@

Table 2: <!--\label{tab:measures}--> Common measures for reporting cancer information.

@@

Introducing population and demographic information helps to interpret the rates in areas effectively, but there will still be uncertainty around the rates. To address this, a cancer atlas often communicates uncertainty about the value of a statistic. There are several potential sources of uncertainty: sampling error, errors arising from the disease reporting process (or data collection), and errors arising from the statistical modelling or simulation process. The most common measures used to present uncertainty are credible or confidence intervals (CIs). Displaying the uncertainty associated with reported statistics is a vital feature of a cancer map, but it is difficult to display effectively. The map focuses on displaying the statistic and lacks additional space to represent the uncertainty. Providing an adjacent map or overlaying maps with symbols [@VSSDCUC] are two common solutions.

## Limitations of choropleth displays {#ch:chorolimit}

@@

<!--
```{r choroCRS, fig.cap = "Four choropleth maps average age-adjusted rate of incidence for lung and bronchus in the United States averaged from sex and years 2012-2016, using four different map projections. The map projections alter the shapes and angles of the boundaries of each state.", message=FALSE, warning=FALSE, fig.align='center'}
#knitr::include_graphics("figures/ggchoroCRS.png")
c1 <- png::readPNG("figures/gg3857.png")
c2 <- png::readPNG("figures/gg2163.png")
c3 <- png::readPNG("figures/gg4326.png")
c4 <- png::readPNG("figures/gg2955.png")
#knitr::include_graphics("figures/usa_grid.png")
load("figures/usa_legend.rda")

@@

```{r ggcartograms, fig.cap = "Figure 2: Common alternatives to maps, showing the same information for the United States of America: (a) contiguous cartogram, (b) non-contiguous, shape-preserved cartogram, (c)  Dorling cartogram (non-contiguous), (d) hexagon tilemap (non-contiguous). Maps (a) - (c) are created by resizing and reshaping the states of the USA to match the 2015 population of the state. This provides a better sense of the extent of disease relative to the population in the country and can help ease losing information about physically small but population-dense states. Map creators give each state equal size and thus equal emphasis in (d) the hexagon tile map.", message=FALSE, warning=FALSE}
knitr::include_graphics("figures/usa_grid.png")
```

Table 3: <!--\label{tab:usa}--> Maps used to present statistics for the United States of America. The colour of each state communicates the average age-adjusted rate of incidence for lung and bronchus for females and males in the United States 2012-2016.

|Map display |Details| 
|:--------------------|:------------------------------------------------|  
|a. Contiguous| It has distorted each state's shape according to the population of the state in 2015. The state of California has become much larger because of the large population density. This draws attention to the densely populated North-East region and detracts from the less populated Mid West.|
|b. Non-contiguous| It maintains the geographic shape of the states, but the size has altered according to the population of the state in 2015. The state of California has remained closer to its original size than its surrounding states. The North-East states have remained closer to their geographical size, for Massachusetts and Connecticut. This draws attention to the densely populated North-East region and the sparse Mid West.|
|c. Dorling| Circles are used to represent each state, but the population of the state determines the size in 2015.  The North-East states remain closer to their neighbors and are slightly displaced from their geographic location. It highlights the sparsity of the population in the Mid West by the distance between the circles at the geographic centroids.|
|d. Hexagon Tessellation| A hexagon of equal size represents each state. It is easy to contrast the neighboring states however the North-East regions have been displaced from their geographic location. It highlights the sparsity of the population in the Mid West by the light yellow color, the Age-Adjusted rate in Kentucky is the darkest and its neighbors are similar.|

@@

To be able to recognize the significant changes, a reader will usually have to know the initial geography to find the differences in the new cartogram layout [@NAC]. The shapes of small areas on a choropleth map and a cartogram are preserved using Tobler's Conformal mapping method.
Koccmoud and House [@CBATCC] present this issue as conflicting tasks or aims, to adjust region sizes and retain region shapes.

### Non-contiguous

@@

## Geofacet
<!-- 166 words -->
Hafen [@IGF] introduces the term geofacet to describe a grid display of small plots. The arrangement of tiles mimics the geographic topology. Geofaceting has the functionality that a statistical plot can be constructed in each facet for each geographic area. A tile map can communicate only one value per region in a visualization, while geofaceting is a more flexible visualization for communication as it increases the amount of information displayed. Virtually any type of plot can be shown in the tile, allowing displays of multiple variables or values per geographic entity. Creating the layout of a geofacet is manual, but once created can be used for any data on that geographic base.

@@

## Multivariate displays
<!-- 222 words -->

@@

# Comparison and critique of alternative displays {#sec:compare}
<!-- 408 words -->

```{r auscartograms, fig.cap = "Figure 4: Cartograms showing melanoma incidence in Australia: (a) contiguous, partially population transformed, (b) non-contiguous shape preserved, (c) Dorling, (d) hexagon tile map. The contiguous cartogram has expanded the highly populated areas while preserving the full shapes of rural areas. If it accurately sized areas by population, the country would be unrecognizable. The shape-preserved is unreadable due to the small area sizes. The Dorling cartogram presents all areas but many are difficult to compare. The hexagon tile map provides a reasonable spatial distribution despite having isolated hexagons in the outback areas.", message=FALSE, warning=FALSE, cache=FALSE, fig.width=8, fig.height=6, dpi=300}
knitr::include_graphics("figures/auscartograms.png")
```

## Neither choropleth maps or cartograms perform well for Australia

@@

Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics (Y=Yes, N=No, S=Sometimes).

@@

# User interaction {#ch:interacting}
<!-- 634 words -->

@@

<!-- ## Under-utilized techniques-->

```{r interacting, fig.cap = "Figure 5: Interactive controls of displays in publicly available choropleth cancer maps: (i) GUI controls for statistic, sex, age groups, continents, and cancer types for [Globocan 2018: Cancer Today ](http://globocan.iarc.fr/Pages/Map.aspx), (ii) Menus for variable selection and zooming on [Bowel Cancer Australia Atlas](http://www.bowelcanceratlas.org/), (iii) Menus for choosing variables and countries in [The Cancer Atlas ](http://canceratlas.cancer.org/data/#?view=map&metric=INCID_ALL_M),  (iv) Tabs for different indicators and cancer types in [Global Cancer Map](http://globalcancermap.com/), (v) Menus and toggles for variable and subset selection in [United States Cancer Statistics: Data Visualizations](https://gis.cdc.gov/Cancer/USCS/DataViz.html).", fig.width = 12, fig.height = 9, results = "asis"}
knitr::include_graphics("figures/interacting.png")
```

@@

```{r animating, fig.cap = "Two examples of advanced interactivity (and animation) in publicly available choropleth cancer maps: a. Linked maps and time-series line plots, with temporal animation in [Map of Cancer Mortality Rates in Spain](http://www.geeitema.org/AtlasET/atlas.jsp?causa=e27MEPOC), b. A highly responsive magnifying glass on a map of [Breast Cancer Mortality in Canada](http://www.ehatlas.ca/light-pollution/maps/breast-cancer-mortality).", fig.width = 12, fig.height = 7, results = "asis"}
knitr::include_graphics("figures/animating.png")
```

# Conclusions
<!-- 307 words -->

<!-- summary of mapping -->
This paper provides an overview of mapping practices as commonly used for cancer atlases and recommends new approaches, such as cartograms and hexagon tile maps that should be adopted going forward. The conventional approach is the choropleth map, and it is widely used. The choropleth map suffers when there are small geographic units, as occurs in Australia where the population is concentrated on the coast, the information about the burden of cancer on those communities can be hidden. Making an inset can clarify congested regions but this breaks the viewers' attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed. The map alternatives implement trade-offs between the familiar shapes, and the importance of the geographic areas in the context of the areas. Given the population or a cancer statistic for each area, the geographic size or shape will change. Alternative displays allow the spatial distribution of cancer data to be digested by map users.
